---
output:
  md_document:
    variant: markdown_github
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
library(sexMat)
```


# sexMat

**Estimate Size at Sexual Maturity**

This package allows to estimate Morphometric and Gonadal Size at Sexual maturity for organisms, usually fish and invertebrates.

The estimation of morphometric maturity used two allometric variables and is based on the relative growth.

The estimation of gonadal maturity used one allometric variable and the stage of sexual maturity (gonad maturation stage).


## Installation

Get the released version from CRAN:

```R
install.packages("sexMat")
```

Or the development version from github:

```R
# install.packages("devtools")
devtools::install_github("ejosymart/sexMat")
```


## Examples

This is a basic example which shows you how to estimate Morphometric and Gonadal Size at Sexual Maturity:

1. Morphometric Size at Sexual Maturity

```{r}
data(crabdata)

classify_data = classify_mature(crabdata, varNames = c("carapace_width", "chela_heigth"), 
                                varSex = "sex_category", selectSex = NULL, method = "ld")

print(classify_data)

# plot(classify_data)

my_ogive = morph_mature(classify_data, method = "fq")

print(my_ogive)

# plot(my_ogive)
```


2. Gonadal Size at Sexual Maturity

```{r}
data(matFish)

my_ogive = gonad_mature(matFish, varNames = c("total_length", "stage_mat"), 
                        inmName = "I", matName = c("II", "III", "IV"), method = "fq", niter = 999)

print(my_ogive)

# plot(my_ogive)
```
